---
title: "mediocrethemes"
description: |
  I created a small package providing not particularly nice and, hopefully, not particularly ugly ggplot themes. Setting the theme for all graphs in a document only requires one short line of code.
categories:
  - R
  - Data viz
  - ggplot
author:
  - name: Vincent Bagilet
    url: {https://vincentbagilet.github.io/vincentbagilet_website}
date: 06-22-2021
citation: false
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, #set to false to hide code
               message = FALSE,
               warning = FALSE,
               out.width = "85%",
               dpi = 300,
               fig.align = "center") 
```

I created a small package providing a ready to use ggplot themes, `mediocrethemes`. The resulting themes are, in my opinion, not strikingly nice but, hopefully, not strikingly ugly either, hence why the name of the package. More info and vignettes to get familiar with the package can be found on the [package website](https://vincentbagilet.github.io/mediocrethemes/).

To me, the most useful feature of my package is the function, `set_mediocre_all()`. When called at the beginning of the code, all subsequent graphs will be created using the new theme, without any additional line of code. This enables users to depart very easily from the nice but too often seen base ggplot themes and color schemes.

I also tried to build my code so that it is reproducible. Users should be able to adapt it to create their own theme if they want to. The code for this package can be find on [GitHub](https://github.com/vincentbagilet/mediocrethemes).

The base color palette is:

```{r}
library(ggplot2)
library(mediocrethemes)

set_mediocre_all()

pal <- unlist(strsplit(colors_table$vector[1], split = ", ")) 

qplot(x = 1:15, y = 1, fill = factor(1:15), geom="tile") +
  theme_void() +
  theme(plot.margin = unit(c(-0.58,-0.9,-0.58,-0.9), "cm")) +
  theme(legend.position="none") 
```


Here are a couple of example plots of the theme and main color scale used but much more can be seen on the [package website](https://vincentbagilet.github.io/mediocrethemes/).

```{r}
ggplot(data = ggplot2::mpg, aes(x = cty, y = displ, fill = manufacturer)) +
    geom_col() +
    labs(title = "A very nice title", subtitle = "A disappointing subtitle")

ggplot(data = ggplot2::mpg, aes(x = cty, y = displ)) +
    geom_point() +
    labs(title = "A very nice title", subtitle = "A disappointing subtitle")
```







